Manuscript Figures

### 2021-09-24 src/manuscript_figures.Rmd
add a panel about LGG clinical feature associated latent factors

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE, message = FALSE, 
                      fig.width = 9, fig.height = 11)
knitr::opts_knit$set(root.dir = params$workdir)
settings <- yaml::read_yaml(params$settings_file)
message(settings$GDCdataDir)
FigureFolder <- params$figures_folder
source(settings$utility_script)
library(data.table)
library(ggplot2)
library(pbapply)
library(ggrepel)
library(ggpubr)
library(cowplot)
library(ggridges)
library(ggcorrplot)
library(ComplexHeatmap)
library(pheatmap)
library(survminer)
library(survival)
library(caret)
library(treeheatr)
library(partykit)

theme_set(ggpubr::theme_pubclean())
geneSetScoresFolder <- '/data/local/buyar/arcas/subtyping_paper/data/geneset_scores/singscore'

@@

require(ggrepel)
dt$label <- stringr::str_wrap(gsub('_', ' ', dt$label), width = 8)
p1 <- ggplot(dt, aes(x = project, y = label)) + 
  geom_point(aes(color = -log10(padj), size = -log10(padj))) + 
  geom_text_repel(aes(label = variable)) + 
  coord_flip() +
  scale_color_gradient(low = 'blue', high = 'red') + 
  labs(x = '', y = 'clinical factor')

@@

dt <- melt(ci, id.vars = 'project')
dt$variable <- ifelse(dt$variable == 'base', 'Age+gender+stage', 'Age+gender+stage + MAUI LFs')

p2 <- ggplot(ci, aes(x = reorder(project, with_lf))) + 
  geom_segment(aes(x = reorder(project, with_lf), xend = project, 
                   y = base, yend = with_lf)) + 
  geom_hline(yintercept = c(0.5, 0.6, 0.7), linetype = 2) + 
  geom_point(data = dt, aes(x = project, y = value, color = variable), size = 2) +
  coord_flip() + 
  labs(x = '', y = '') + 
  theme(legend.title = element_blank(), legend.position = 'bottom', 
        legend.background = element_blank(), 
        legend.margin = margin(), 
        legend.direction = 'vertical', legend.box.background = element_blank()) +
  scale_color_brewer(type = 'qual', palette = 6)

Fig <- cowplot::plot_grid(cowplot::plot_grid(p1, p2, nrow = 2, labels = c('A', 'B')))

fig_count <- fig_count + 1
Figures[[fig_count]] <- Fig
fig.name <- paste0('Figure', fig_count, '.clinical_factors.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(filename = file.path(FigureFolder, fig.name), 
       plot = Fig, width = 8.27, height = 11.69, units = 'in')
Legends[[fig_count]] <- "A) Top significant (padj < 0.05) clinical factor-associated multi-omic fingerprints per cancer type. B) Harrell’s C-index for progression-free survival prediction accuracy computed using only clinical factors (age + gender + tumor stage) or using clinical factors along with survival-predictive MAUI fingerprints"

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE, message = FALSE, 
                      fig.width = 9, fig.height = 11)
knitr::opts_knit$set(root.dir = params$workdir)
settings <- yaml::read_yaml(params$settings_file)
message(settings$GDCdataDir)
FigureFolder <- params$figures_folder
source(settings$utility_script)
library(data.table)
library(ggplot2)
library(pbapply)
library(ggrepel)
library(ggpubr)
library(cowplot)
library(ggridges)
library(ggcorrplot)
library(ComplexHeatmap)
library(pheatmap)
library(survminer)
library(survival)
library(caret)
library(treeheatr)
library(partykit)
theme_set(ggpubr::theme_pubclean())

geneSetScoresFolder <- '/data/local/buyar/arcas/subtyping_paper/data/geneset_scores/singscore'

@@

require(ggrepel)
dt$label <- stringr::str_wrap(gsub('_', ' ', dt$label), width = 8)
p1 <- ggplot(dt, aes(x = project, y = label)) + 
  geom_point(aes(color = -log10(padj), size = -log10(padj))) + 
  geom_text_repel(aes(label = variable)) + 
  coord_flip() +
  scale_color_gradient(low = 'blue', high = 'red') + 
  theme(axis.title = element_blank())

# make example plots for selected clinical features
pr <- 'TCGA-LGG' 
datatype <- 'cnv_gex_meth_mut'
M <- LFs$maui_factors[[paste0(pr, '.', datatype)]]
clin <- process_clinical_covariates(rownames(M))
dt.pr <- dt[padj < 0.05][project == pr][order(padj)]
# select top 3 clinical features for the project
plots <- lapply(head(dt.pr$label, 3), function(x) {
  # get lf, clinical feature and plot distribution 
  f <- gsub("[\n ]", "_", x)
  lf <- dt.pr[label == x]$variable
  df <- cbind(data.frame(M[,lf,drop = F]), clin[,f,drop=F])
  df <- df[!is.na(df[,f]),]
  ggpubr::ggboxplot(df, y = lf, x = f, 
                    color = f, add = 'jitter') + 
    labs(x = x) + 
    theme(legend.position = 'none', axis.text.x = element_text(angle = 45, hjust = 1)) 
})

p2 <- cowplot::plot_grid(plotlist = plots, nrow = 1, align = 'h')

@@

dt <- melt(ci, id.vars = 'project')
dt$variable <- ifelse(dt$variable == 'base', 'Age+gender+stage', 'Age+gender+stage + MAUI fingerprints')

p3 <- ggplot(ci, aes(x = reorder(project, with_lf))) + 
  geom_segment(aes(x = reorder(project, with_lf), xend = project, 
                   y = base, yend = with_lf)) + 
  geom_hline(yintercept = c(0.5, 0.6, 0.7), linetype = 2) + 
  geom_point(data = dt, aes(x = project, y = value, color = variable), size = 2) +
  coord_flip() + 
  labs(x = '', y = '') + 
  theme(legend.title = element_blank(), legend.position = 'bottom', 
        legend.background = element_blank(), 
        legend.margin = margin(), 
        legend.direction = 'vertical', legend.box.background = element_blank()) +
  scale_color_brewer(type = 'qual', palette = 6)

Fig <- cowplot::plot_grid(p1, 
                          cowplot::plot_grid(p2, p3, 
                                             nrow = 1, 
                                             labels = c('B', 'C'), 
                                             rel_widths = c(2, 2)), 
                          labels = c('A', ''), 
                          nrow = 2)

fig_count <- fig_count + 1
Figures[[fig_count]] <- Fig
fig.name <- paste0('Figure', fig_count, '.clinical_factors.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(filename = file.path(FigureFolder, fig.name), 
       plot = Fig, width = 8.27, height = 11.69, units = 'in')
Legends[[fig_count]] <- "A) Top significant (padj < 0.05) clinical factor-associated multi-omic fingerprints per cancer type. B) MAUI fingerprints enriched for subgroups of different clinical features of TCGA-LGG samples. 
LF45 is enriched among samples histologically categorized as Oligodendrogliomas; LF54 is enriched among donors of male gender; LF65 is enriched among samples with histological grade of G3. C) Harrell’s C-index for progression-free survival prediction accuracy computed using only clinical factors (age + gender + tumor stage) or using clinical factors along with survival-predictive MAUI fingerprints"
